Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-09-30 manuscript/manuscript.tex
edited disco

Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
that diversity is reduced at linked neutral variation as an advantageous
polymorphism sweeps to fixation, over four decades of theoretical and empirical
research has enriched our understanding of linked selection.

An alternative approach to understanding the
genome-wide effects of selection on standing variation, e.g. selection on an
infinitesimal polygenic trait, stems from an early quantitative genetic model
of linked selection \parencite{Robertson1961-ho} and its later developments
(\cite{Santiago1995-hx,Santiago1998-bs}; see also \cite{Barton2000-zg} for a
comparison of these models with classic hitchhiking models). Implicit in these
models is that autocovariance between allele frequency change is created when
there is heritable fitness variation in the population, a signal that can be
readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
we provide the first empirical evidence of linked selection acting on standing
variation over very short timescales.

It is worth building some intuition why temporal covariance allows us to detect
such faint signals of polygenic linked selection from temporal genomic data.
Each variant is subject to both variance in allele frequency due to drift and
two levels of sampling noise (as individuals are sampled from the population
and reads are sampled from amplified DNA), which at any locus swamps the
temporal covariance signal and creates spurious covariances. However, these
spurious covariances do not share a common sign across timepoints whereas the
covariances created by linked selection do; consequently, averaging across the
entire genome, the temporal signal exceeds sampling noise.

Depending on how many loci affect fitness, such a strong
effect of linked selection may not be differentiable from genetic drift using
only single contemporary population samples. In this way, temporal data allows
us to sidestep the key problem of detecting selection from standing variation:
that the genomic footprint leaves too soft of a signature to differentiate from
a background of genetic drift. In fact we find that the temporal covariance
signal is detectable even in the most extremely difficult to detect soft sweep
case: polygenic selection on an infinitesimal trait.

@@

We reanalyzed three published temporal genomic datasets and detected the
genome-wide signature of linked selection acting over tens of generations.
Furthermore, we find the dynamics of temporal autocovariance are consistent
with theory; the temporal covariance in allele frequencies decays as both
the fitness variation associated with particular haplotypes and the linkage
disequilibria between selected and neutral loci decay through time. In
reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
generations, greater than 20\% of the variation in allele frequency change is
directly due to the action linked selection. Capitalizing on replicated
evolve-and-resequence study designs, we characterized the extent to which
convergent selection pressures lead to parallel changes in allele frequencies
across replicate populations, finding moderate correlations in allele frequency
changes.

While we must keep in mind that the studies analyzed here were all laboratory
populations and natural selection in the wild will likely differ, overall we
have found that linked selection has a remarkably strong effect for such short
study durations.

One limitation of these analyses is that none of the studies we reanalyzed
estimated linkage disequilibria data for the evolved populations. Our theory of
temporal autocovariance tells us that the temporal autocovariances a neutral
site experiences is determined by the product of the expected linkage
disequilibrium and additive fitness variation that persists through the
generations \parencite{Buffalo2019-io}. This leads to a clear prediction:
regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. Unfortunately, we lack a high-resolution recombination map for
\emph{D. simulans} and while there are LD data for \emph{D. simulans}
\parencite{Signor2018-wg} we did not find a relationship between temporal
covariance and LD.

% \graham{TODO Is important point. We'll likely need plots for supp}

We believe this is driven by the idiosyncratic nature of LD in
evolve-and-resequence populations \parencite{Nuzhdin2013-gf,Kelly2019-dc}, and
that we might find such a relationship if LD data from the evolved population
were available. Unfortunately, sequencing multiple timepoints is expensive and
the low-coverage and/or pooled-sequencing approaches common in these studies
prohibit estimating linkage disequilibrium. Future studies complete with LD
data and recombination maps would allow one to disentangle the influence of
closely linked sites from more distant sites in causing temporal
autocovariance, and allow us to better understand whether sites in high
recombination regions are free from the effects of linked selection.
Furthermore, such additional data would allow for localizing the effects of
temporal covariance, which would allow us to estimate not just the genome-wide
fraction of variation in frequency change ($G(t)$), but also the fraction of
the genome impacted by linked selection, allowing us to synthesize the temporal
approach with single-timepoint studies like that of \textcite{Elyashiv2016-vt}.

%Why the variation in convergence correlations?

Thus far, the most comprehensive studies implicating linked selection in
affecting genome-wide diversity have been in \emph{Drosophila}
\parencite{Begun1992-ey,Elyashiv2016-vt,Sattath2011-dr} and
\emph{Caenorhabditis} \parencite{Cutter2003-tl,Cutter2003-tl,Andersen2012-bj},
whereas the evidence for linked selection in plant and yeast species is weak
\parencite{Cutter2013-ba}. Both \emph{Drosophila} and \emph{Caenorhabditis}
have short genetic maps (or in the case of selfing \emph{C.  elegans}, small
effective recombination rates), and large effect population sizes, two factors
predicted to increase the degree to which hitchhiking impacts genome-wide
diversity \parencite{Barton2000-zg}. However, our results here show that even
with small effective population sizes (300, 450, and 45 for the
\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies respectively)
and, in the case of mice, moderately-sized genetic maps (around 14 Morgans;
\cite{Cox2009-hf}), we still see a fairly strong effect of linked selection.
This suggests that while these factors may govern the dynamics of classic
hitchhiking events, they could perhaps play less of a role in polygenic linked
selection. However, with only three cases analyzed, further studies on
different taxa are needed.

Finally, in reanalyzing the \textcite{Barghi2019-qy} study, we find evidence
of complex linked selection dynamics. This is due either to a change in the
environment and its consequences for the fitness of regions in the genome,
recombination breaking up selected alleles, or epistatic interactions between
loci. It is important to note that the original study did not intentionally
alter the environment in any way; this signal is attained without intentionally
seeking it out. Discerning which of these is creating negative temporal
autocovariance is a sizable challenge, requiring sequencing each generation and
LD data.

Overall, we hope this will encourage more temporal genomic studies in both
laboratory and natural populations. Understanding the dynamics of linked
selection over short timescales will help to unite phenotypic studies of rapid
adaptation of polygenic adaptations with a detectable genomic signature, to
address long-standing questions concerning linked selection, evolutionary quantitative
genetics, and the overall impact selection has on genetic variation.

>>>>

Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
that linked neutral diversity is reduced as an advantageous
polymorphism arises and sweeps to fixation, over four decades of theoretical and empirical
research has enriched our understanding of linked selection. 
One under-used approach to understand the
genome-wide effects of selection on standing variation, e.g. selection on an
infinitesimal polygenic trait, stems from an early quantitative genetic model
of linked selection \parencite{Robertson1961-ho} and its later developments
(\cite{Santiago1995-hx,Santiago1998-bs}; see also \cite{Barton2000-zg} for a
comparison of these models with classic hitchhiking models). Implicit in these
models is that autocovariance between allele frequency change is created when
there is heritable fitness variation in the population, a signal that may be
readily detected from temporal genomic data \parencite{Buffalo2019-io}. 
Depending on how many loci affect fitness, such a strong
effect of linked selection may not be differentiable from genetic
drift using only single contemporary population samples or looking at
at temporal allele frequency change at each locus in isolation.  In
this way, averaging summaries of temporal data allows us to sidestep the key problem of detecting selection from standing variation:
that the genomic footprint leaves too soft of a signature to differentiate from
a background of genetic drift. In fact we find that the temporal covariance
signal is detectable even in the most extremely difficult to detect soft sweep
case: polygenic selection on highly polygenic traits \parencite{Buffalo2019-io}.

It is worth building some intuition why temporal covariance allows us to detect
such faint signals of polygenic linked selection from temporal genomic data.
Each variant is subject to both variance in allele frequency due to drift and
sampling noise, which at any locus may swamp the temporal covariance signal and creates spurious covariances. However, these
spurious covariances do not share a directional signal whereas the
covariances created by linked selection do; consequently, averaging across the
entire genome, the temporal signal exceeds sampling noise.

@@

Our analyses reveal that a sizable proportion of allele frequency
change in these populations is due to the (indirect) action of
selection. Capitalizing on replicated designs, we characterized the extent to which
convergent selection pressures lead to parallel changes in allele frequencies
across replicate populations, and found that a reasonably proportion
of the response is shared across short timescales. These likely
represent substantial under-estimates of the contribution of linked
selection because ....

These estimates of the contribution of selection could be refined by using patterns of LD and
recombination which would allow us to more fully parameterize a linked-selection
model of temporal allele frequency change \parencite{Buffalo2019-io}. The basic prediction is that regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. However, one limitation of these pooled sequence datasets is that none of the studies we reanalyzed
estimated linkage disequilibria data for the evolved
populations. There are LD data for a natural
population of \emph{D. simulans} \parencite{Signor2018-wg},  we did not find a relationship between temporal
covariance and LD.  We believe this is driven by the idiosyncratic nature of LD in
evolve-and-resequence populations, which often extends over large
genomic distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD
data and recombination maps would allow one to disentangle the influence of
closely linked sites from more distant sites in causing temporal
autocovariance, and allow the fitting of more parameteric model to
estimate population parameters such as the additive genetic variation
for fitness directly from temporal gneomic data alone  \parencite{Buffalo2019-io}.

Our primary focus here has been on evolution in laboratory
populations. It is unclear whether we should expect a similar impact
of selection in natural populations. In some of these experiments, selection pressures may
have been stronger or more sustained that in natural populations REF.
Conversely, they these lab populations were maintained at very small
population sizes, small effective population sizes 300, 450, and 45 for the
\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies
respectively, which will amplify the role of genetic drift. The
advantage of lab experiments is that they are closed populations, in
natural populations temporal covariance could also arise from
the systematic migration of alleles from differentiated
populations. To these methods to natural populations will require
either populations that are reasonably closed to migration, or for the effect of migration to
be accounted for possibly either by knowledge of allele frequencies in source
populations or the identification of migrant individuals.

While it challenging to apply temporal methods to natural populations
there is a lot of promise for these approaches \cite{Dmitri_others}. 
Efforts to estimate the quantify the impact of linked selection in
humans and {\it Drosophila melanogaster} have suggested that XXX to
XXX of allele frequency change over long time periods is driven by
selection \citep{}. Thus there is reason to be hopeful that we could
detect this from temporal data, which would help to resolve the
time-scales that linked selection act over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
of complex linked selection dynamics, with selection pressures
flipping over time due to either environmental change, the breakup of
epistatic combinations or advantagous haplotypes. Such patterns would
 be completely obscured in samples only from contempory populations,
 thus we can hope tyo have a much richer picture of linked selection
 as temporal sequencing becomes more common.

%Overall, we hope this will encourage more temporal genomic studies in both
%laboratory and natural populations.

Understanding the dynamics of linked
selection over short timescales will help to unite phenotypic studies of rapid
adaptation with a detectable genomic signature, to
address long-standing questions concerning linked selection, evolutionary quantitative
genetics, and the overall impact selection has on genetic variation.
